Representing Scenes as Light Fields for Real Time Rendering in Virtual Reality Display

### 2021-03-09 TOG/5-result.tex
submission

\section{Evaluation}
\label{sec:result}
\input{TOG/fig_tile_results}
With a variety of scenes as shown in \Cref{fig:results:comparison}, we first conduct subjective studies (\Cref{sec:study:user}) and objective measurements (\Cref{sec:study:quality}) to evaluate the perceptual quality.
Further, we analyze the intra-system workings and compare inter-system performance in \Cref{sec:study:intra}.

\subsection{User Study}
\label{sec:study:user}
% \zh{(Jan 10) all the Qs are answered offline, will update the section soon.}
% In this section, we report two perceptual experiments to investigate 1) how users perceive our solution and the ground truth and 2) how our solutions behaves compared to other alternatives.
We first conducted a psychophysical experiment to investigate how users perceive our solution ({\bf OURS}) compared with existing neural view synthesis methods (\cite{mildenhall2020nerf}, {\bf NeRF}) and foveated rendering with the full mesh data (\cite{perry2002gaze}, {\bf F-GT}).

@@

\paragraph{Stimuli}
For precise comparison and accommodating the low frame rate of alternative solutions (as to be compared in \Cref{sec:study:intra}), each group of stimuli were static images rendered via {\bf F-GT} or synthesized via  {\bf OURS}/{\bf NeRF} with the same views, as shown in \Cref{fig:results:comparison}.
They were generated with the same and randomly defined gaze fixation across conditions.
The resolution of the image per eye was $1440 \times 1600$. 
During the study, the target gaze position was indicated as a green cross on the stimuli images.
% The resolution of the images per eye was $1440 \times 1600$. 
We used the {\it gas} and {\it minecraft} scenes for the diverse indoor/outdoor environment.
Each condition from an individual scene consists of \nothing{$1$ view and} $3$ different gaze positions.

The foveal kernel size in {\bf F-GT} was designed to match our display size and eye-display distances.
For generating {\bf NeRF} condition, we retrained the model from \cite{mildenhall2020nerf} on our dataset with $N_{sample}=16$, $N_{layers}=4$, $N_{channel}=128$. The bottom right of each scene figure \Cref{fig:results:comparison} shows the resulting images. 
%We predict the image from the camera transformation, fine-tune the foveated layer, and render the image by alpha blending the three layers.

@@

\paragraph{Setup}
Each participant wore an eye-tracked HTC Vive Pro Eye headset and remained seated to example the stimuli during the experiment.
Ten users participated in and completed the study (3 females, $M=23.7$, $SD=1.49$). 
% One participant (age=23, female) did not complete the experiment due to constant gaze drifting (please refer to the task description below).
None of the subjects were aware of the research, the experimental hypothesis, nor the number of methods. All participants had a normal or corrected-to-normal vision.

\paragraph{Task}
The task was a \textit{two-alternative-forced-choice} (2AFC). Each trial consists of a pair of stimuli generated from two of the three methods ({\bf OURS} / {\bf NeRF} / {\bf F-GT}) with a sampled view and gaze position. Each stimulus appeared for $300$ms on display. A forced $1.5$sec break (black screen) was introduced between conditions to clear the vision. During the study, the participants were instructed to fix their gazes on the green cross. To prevent the fixation from shifting away and ensure accuracy, we tracked the users' gaze throughout the experiment. Whenever the gaze is more than $5$deg away from the target, a trial was dropped immediately with a black screen informing the participant.
After each trial, the participants were instructed to select which of the two stimuli appeared with higher visual quality using a keyboard.
Before each experiment, a warm-up session with 6 trials was provided to familiarize the participants with the study procedure. The orders of conditions among trials were randomized with counterbalancing. The entire experiment of each user consists of $36$ trials, $12$ trial per pair of conditions.
%To start with, each participant was informed that they were required to choose an alternative with higher quality after watching each pair of stimuli. We calibrated the device and performed a warm-up trial to help participants understand our system. Each participant were provided with 1 warm-up trial and 6 official trials (2 scenes x 1 view x 3 gaze positions) where each trial consisted all the combinations of the pairs of every two stimuli. We randomize the order of the pairs for counterbalancing. We showed a yellow cross on the screen. After participants move their gaze positions to the area close to the cross, the yellow cross turns green. The first stimulus will be rendered as well as the cross after participants fixate the cross for 1 second. Each stimulus is displayed for 300 milliseconds. The second one will be rendered after 1.5-second dark screen. If the participants does not fixate at the cross when stimuli are displayed, the current pair will be discarded and displayed to the participants later in the same trial. After each pair, participants reported their opinions on which image has higher quality.
To minimize the effect of accumulated fatigue, we enforced breaks at least a 2-second between trials and 60-second after each scene. Meanwhile, the participants were able to take as much time as needed.

\paragraph{Results}
\input{TOG/fig_vote}
\Cref{fig:results:2afc} plots the results considering both methods and their orders in the 2AFC experiments. Among all three conditions, we observed close-to-random-guess among trails that compares {\bf F-GT} and {\bf OURS} ($53.3\%$ voted for {\bf OURS}, $SD=14.9\%$). Meanwhile, a significantly higher ratio of voting {\bf F-GT} over {\bf NeRF} was observed ($92.5\%$, voted for {\bf F-GT}, $SD=11.4\%$, binomial test showed $p<0.005^{***}$). The preference applies to {\bf OURS} vs. {\bf NeRF} as well ($91.7\%$ voted for {\bf OURS}, $SD=14.8\%$, binomial test showed $p<0.005^{***}$).
% \qisun{(Jan 23, 2021) @Zhenyi: consider adding stdev to the average numbers above? This shall be reflected in \Cref{fig:results:2afc} as well. Also, double check the numbers between here and the plot as they are not matching now.}

@@

Meanwhile, both of the two conditions showed significant quality preference than {\bf NeRF}. That is, with immersive, high FoV, and first-person viewing constraints, {\bf OURS} synthesizes gaze-contingent retinal image with superior perceptual quality than alternative synthesis solutions. The latter typically aims to reconstruct the full image over the entire field-of-view. \nothing{the current volume-based scene representation may not fully reproduce the perceptually identical retinal images.}
To further validate the perceptual quality at individual eccentricities across the whole visual field, we analyze with objective metrics in the following section.

\subsection{Visual Quality}
\label{sec:study:quality}
\input{TOG/fig_lpips}
Complementary to the subjective measurement (\Cref{sec:study:user}), we evaluate the perceived quality in an objective fashion. Specifically, under all four different scenes (indoor/outdoor, artificial/realistic), we compare {\bf OURS}, {\bf F-GT} and {\bf NeRF} at individual eccentricity ranges (up to $110$ deg, the capability of the VR HMD) via partial images. 
For each eccentricity range, we compare the deep perceptual similarity (LPIPS) \cite{zhang2018unreasonable} across all scenes. LPIPS uses deep neural networks to estimate perceptual similarities between the image provided and a reference image. Smaller values indicate higher perceptual similarity.
For each scene, we sample $20$ views with gazes at the middle of the display, resulting in $20$ data per eccentricity value, $420$ data per scene. We used one-way repeated measures ANOVAs to compare effects across three stimuli on each eccentricity range and together. Paired t-tests with Holm correction were used for all pairwise comparisons between stimuli. All tests for significance were made at the $\alpha=0.05$ level. 
% The error bars in the graphs show the 95\% confidence intervals of the means.

\paragraph{Results} 
% \warning{need to update}
\Cref{fig:lpips} plots LPIPS values across all scenes and eccentricity ranges ($5$ deg step size). 
From foveal to near eccentricity (<=$40$ deg), we observed significant effects of the stimuli on LPIPS with a ``large'' effect size ($\eta^2 = 0.26$). That is, \textbf{OURS} shows significantly lower LPIPS than \textbf{NeRF} ($p<.001^{***}$), although higher than \textbf{F-GT} ($p<.001^{***}$).
For example, the main effects of stimuli ($F(2,38)=210.76, p<.001^{***}$) was significant on eccentricity $=[0,25]$ deg in scene \textit{gallery} (\Cref{fig:lpips:gallery}). \textbf{OURS} was significantly lower than \textbf{NeRF} ($t(19)=-19.54, p<.001^{***}$) and \textbf{F-GT} ($t(19)=-4.32, p<.001^{***}$) both with a ``large'' effect size (Cohen's d $>0.8$). The example observation generally applies to all 4 scenes being validated.

From mid- to far- periphery (>$40$ deg), we observed significant effects of the stimuli on LPIPS with a ``large'' effect size ($\omega^2 = 0.16$). \textbf{OURS} shows significantly higher LPIPS than \textbf{NeRF}. Whereas, comparing with \textbf{F-GT}, we observed significant lower scores ($p<.001^{***}$).
For instance, the main effects of stimuli was significant on eccentricity $=60$ deg in scene \textit{bedroom} ($F(2,38)=411.93, p<.001^{***}$, \Cref{fig:lpips:bedroom}). \textbf{OURS} was significantly lower than \textbf{F-GT} ($t(19)=-3.78, p<.001^{***}$), and higher than \textbf{NeRF} ($t(19)=22.75, p<.001^{***}$) both with a ``large'' effect size (Cohen's d $>0.8$). The example observation generally applies to all 4 scenes being validated.

\paragraph{Discussion}
In the fovea and near-periphery, the observation revealed our method's significantly higher perceptual quality than alternative neural representation solutions. This is evidenced by the significantly stronger perceptual similarity to \textbf{F-GT} by comparing between \textbf{OURS} and \textbf{NeRF}. The subjective study (\Cref{sec:study:user}) evidenced the marginally lower similarity than \textbf{F-GT} as statistically unnoticeable among users.

In the far-periphery, \textbf{OURS} showed increased LIPIPS than \textbf{F-GT} beyond $40$ deg.
The latter has been shown to display identical perceptual similarity than full resolution rendering in VR \cite{Patney:2016:TFR}.
Thus, the findings show that \textbf{OURS} doesn't compromise the peripheral vision's quality with its significantly enhanced synthesis acuity in fovea.

Note that the discoveries also agree with the observations from \Cref{sec:study:user}. That is, in addition to the significantly faster performance  ($99.7\%$ time reduction per frame as in \Cref{sec:study:intra}\nothing{\qisun{Check if I noted it right.}}), our method showed superior perceptual quality than \textbf{NeRF} under first-person, high resolution, and immersive viewing of 3D scenes. Meanwhile, our synthesized views are perceptually identical to the traditionally rendered images with a complete 3D mesh. It yet achieves significant data storage-saving with only about $0.5$mb (up to $99.5\%$ in our experiment).

\subsection{Performance}
\label{sec:study:intra}
% apply objective metric to |cref{fig:optimization}, aka different sample numbers
Virtual and augmented reality demands high frame rates along with high quality to ensure immersive and comfortable experience. Our neural synthesis method introduces both spatial and angular (stereo) visual perception for real-time performance (\Cref{sec:method:representation}). 
Further, our precision-latency joint optimization for the representation and network also balances quality and performance.
Here, we evaluate the performance of each component in our system and compare with existing neural synthesis solutions.

For high resolution ($1440\times1600$), high field-of-view ($110$deg) stereo images required by VR display, our system completes all computation (including gaze-contingent neural-inference and elemental images composition) in $31.8$ms per frame without stereo foveation, \dncR{as shown in \Cref{tbl:ablation} \st{shows a time consumption breakdown of our method, together with {\bf NeRF}}}. Compared with {\bf NeRF} \dncR{(the bottom row of \Cref{tbl:ablation})}, our spatial foveation \dncR{\st{for adaptive resolution}} significantly improves the system performance from offline computation (in seconds) to interactive speed (\dncR{about 30 frames per second\st{$31.8$ ms per frame}}).

Tailored for VR displays, our \dncR{stereo foveation\st{adaptive stereo synthesis}} further reduces the computation time \dncR{and achieves high performance of more than 50 frames per second \st{by more than $10$ms/frame}}, contributing to a temporally continuous viewing experience without loss of perceived resolution and quality.

%Specifically, the significant performance gain benefits from the spatial and stereo foveation that do not compromise the visual quality.
\begin{table}[!htb]
    \centering
    \begin{tabular}{ l|r } 
        \toprule
         foveal infer (per eye)  & 3.8 \\
         \midrule
         periphery infer (per eye) & 11.9 \\
          \midrule
         blending \& contrast enhancement & 0.2 \\
          \midrule\midrule
         {\bf OUR} method without stereo foveation & 31.8\\
         {\bf OUR} full method & {\bf 19.7}\\
         \midrule
         {\bf NeRF} \cite{mildenhall2020nerf} & $9\times10^3$ \\
         \bottomrule
    \end{tabular}
    \Caption{Time consumption breakdown and comparison.}
    {%
    The numbers show the time consumption of each component and the overall system per frame. All units are in millisecond.
    \nothing{\qisun{(Jan 25) I have update the time consumption of {\bf NeRF}. See if that is correct.}}
    }
    \label{tbl:ablation}
\end{table}

>>>>

\section{Evaluation}
\label{sec:result}
% \input{TOG/fig_tile_results}
With a variety of scenes as shown in \Cref{fig:results:comparison}, we conduct a subjective study (\Cref{sec:study:user}), and calculate objective measurements (\Cref{sec:study:quality}) to evaluate the perceptual quality.
Further, we analyze the intra- and inter-system performance in \Cref{sec:study:intra}.

\subsection{User Study}
\label{sec:study:user}
% In this section, we report two perceptual experiments to investigate 1) how users perceive our solution and the ground truth and 2) how our solutions behaves compared to other alternatives.
We conducted a psychophysical experiment to investigate how users perceive our solution ({\bf OURS}) compared with existing neural view synthesis methods (\cite{mildenhall2020nerf}, {\bf NeRF}) and foveation rendering of the original mesh scene (\cite{perry2002gaze}, {\bf F-GT}).

@@

\paragraph{Stimuli}
For precise comparison and accommodating the low frame rate of alternative solutions (as to be compared in \Cref{sec:study:intra}), each group of stimuli consisted of static stereo images rendered via {\bf F-GT} or synthesized via {\bf OURS}/{\bf NeRF} with the same views, as shown in \Cref{fig:results:comparison}.
They were generated with the same and randomly defined gaze fixation across conditions.
The resolution of the image per eye was $1440 \times 1600$. 
During the study, the target gaze position was indicated as a green cross on the stimuli images.
% The resolution of the images per eye was $1440 \times 1600$. 
We used the {\it gas} and {\it minecraft} scenes for the diverse indoor/outdoor environment.
Each condition from an individual scene consists of \nothing{$1$ view and} $3$ different gaze positions.

The foveal kernel size in {\bf F-GT} was designed to match our display size and eye-display distances.
For generating {\bf NeRF} condition, we retrained the model from \cite{mildenhall2020nerf} on our dataset with $\mlpLayerNum=4$, $\mlpChannelNum=128$, and $\sphereNum = 16$. The bottom right of each scene figure \Cref{fig:results:comparison} shows the resulting {\bf NeRF} images. 
%We predict the image from the camera transformation, fine-tune the foveated layer, and render the image by alpha blending the three layers.

@@

\paragraph{Setup}
Each participant wore an eye-tracked HTC Vive Pro Eye headset and remained seated to examine the stimuli during the experiment.
Ten users participated in and completed the study (3 females, $M=23.7$, $SD=1.49$). 
% One participant (age=23, female) did not complete the experiment due to constant gaze drifting (please refer to the task description below).
No participants were aware of the research, the experimental hypothesis, nor the number of methods. All participants had a normal or corrected-to-normal vision.

\paragraph{Task}
The task was a \textit{two-alternative-forced-choice} (2AFC). Each trial consists of a pair of stimuli generated from two of the three methods ({\bf OURS} / {\bf NeRF} / {\bf F-GT}) with a sampled view and gaze position. Each stimulus appeared for $300$ms on display. A forced $1.5$sec break (black screen) was introduced between conditions to clear the vision. During the study, the participants were instructed to fix their gazes on the green cross. To prevent the fixation from shifting away and ensure accuracy, we tracked the users' gaze throughout the experiment. Whenever the gaze is more than $5$ deg away from the target, a trial was dropped immediately with a black screen informing the participant.
After each trial, the participants were instructed to select which of the two stimuli appeared with higher visual quality using a keyboard.
Before each experiment, a warm-up session with 6 trials was provided to familiarize the participants with the study procedure. The orders of conditions among trials were randomized and counter-balanced. The entire experiment of each user consists of $36$ trials, $6$ trials per pair of ordered conditions.
% \zh{Regarding ``$12$ trial per pair of conditions'', here you treat pair(a-b) and pair(b-a) the same?}\qisun{I did this before knowing your plot on treating them differently, which was better. Tweaked.}
%To start with, each participant was informed that they were required to choose an alternative with higher quality after watching each pair of stimuli. We calibrated the device and performed a warm-up trial to help participants understand our system. Each participant were provided with 1 warm-up trial and 6 official trials (2 scenes x 1 view x 3 gaze positions) where each trial consisted all the combinations of the pairs of every two stimuli. We randomize the order of the pairs for counterbalancing. We showed a yellow cross on the screen. After participants move their gaze positions to the area close to the cross, the yellow cross turns green. The first stimulus will be rendered as well as the cross after participants fixate the cross for 1 second. Each stimulus is displayed for 300 milliseconds. The second one will be rendered after 1.5-second dark screen. If the participants does not fixate at the cross when stimuli are displayed, the current pair will be discarded and displayed to the participants later in the same trial. After each pair, participants reported their opinions on which image has higher quality.
To minimize the effect of accumulated fatigue, we also enforced breaks between trials (at least 2 seconds) and after each scene (60 seconds). Meanwhile, the participants were allowed to take as much time as needed.

\paragraph{Results}
\input{TOG/fig_vote}
\Cref{fig:results:2afc} plots the results considering both methods and their orders in the 2AFC experiments. Here we analyzed and reported the results regardless of the orders. Among all three conditions, we observed close-to-random-guess among trials that compared {\bf F-GT} and {\bf OURS} ($53.3\%$ voted for {\bf OURS}, $SD=14.9\%$). Meanwhile, a significantly higher ratio of voting {\bf F-GT} over {\bf NeRF} was observed ($92.5\%$, voted for {\bf F-GT}, $SD=11.4\%$, binomial test showed $p<0.005^{***}$). The preference applies to {\bf OURS} vs. {\bf NeRF} as well ($91.7\%$ voted for {\bf OURS}, $SD=14.8\%$, binomial test showed $p<0.005^{***}$).
% \qisun{(Jan 23, 2021) @Zhenyi: consider adding stdev to the average numbers above? This shall be reflected in \Cref{fig:results:2afc} as well. Also, double check the numbers between here and the plot as they are not matching now.}

@@

Meanwhile, both of the two conditions showed significant quality preference than {\bf NeRF}. That is, with immersive, high FoV, and first-person viewing constraints, {\bf OURS} synthesizes gaze-contingent retinal images with superior perceptual quality than alternative synthesis solutions. The latter typically aims to reconstruct the full image over the entire field-of-view. \nothing{the current volume-based scene representation may not fully reproduce the perceptually identical retinal images.}
To further validate the perceptual quality at individual eccentricities across the whole visual field, we analyze with objective metrics in the following section.

\subsection{Visual Quality}
\label{sec:study:quality}
\input{TOG/fig_lpips}
Complementary to the subjective measurement (\Cref{sec:study:user}), we evaluate the perceived quality in an objective fashion. Specifically, under all four different scenes (indoor/outdoor, artificial/realistic), we compare {\bf OURS}, {\bf F-GT}, and {\bf NeRF} at individual eccentricity ranges (up to $110$ deg, the capability of the VR HMD) via partial images. 
For each eccentricity range, we compare the deep perceptual similarity (LPIPS) \cite{zhang2018unreasonable} across all scenes. LPIPS uses deep neural networks to estimate perceptual similarities between the image provided and a reference image. Smaller values indicate higher perceptual similarity.
For each scene, we sample $20$ views with gazes at the middle of the display, resulting in $20$ data per eccentricity value, $21$ eccentricity value per scene ($5$ deg step size), thus $420$ data per scene. We used one-way repeated measures ANOVAs to compare effects across three stimuli on each eccentricity range and together. Paired t-tests with Holm correction were used for all pairwise comparisons between stimuli. All tests for significance were made at the $\alpha=0.05$ level. 
% The error bars in the graphs show the 95\% confidence intervals of the means.

\paragraph{Results} 
% \warning{need to update}
\Cref{fig:lpips} plots LPIPS values across all scenes and eccentricity ranges ($5$ deg step size). 
From foveal to near eccentricity ($\leq40$ deg), we observed significant effects of the stimuli on LPIPS with a ``large'' effect size ($\eta^2 = 0.26$). That is, \textbf{OURS} shows significantly lower LPIPS than \textbf{NeRF} ($p<.001^{***}$), although higher than \textbf{F-GT} ($p<.001^{***}$).
For example, the main effects of stimuli ($F(2,38)=210.76, p<.001^{***}$) was significant on eccentricity $=[0,25]$ deg in scene \textit{gallery} (\Cref{fig:lpips:gallery}). \textbf{OURS} was significantly lower than \textbf{NeRF} ($t(19)=-19.54, p<.001^{***}$) and \textbf{F-GT} ($t(19)=-4.32, p<.001^{***}$) both with a ``large'' effect size (Cohen's d $>0.8$). The example observation generally applies to all 4 scenes being validated.

From near- to far- eccentricity (>$40$ deg), we observed significant effects of the stimuli on LPIPS with a ``large'' effect size ($\omega^2 = 0.16$). \textbf{OURS} shows significantly higher LPIPS than \textbf{NeRF}. Whereas, comparing with \textbf{F-GT}, we observed significant lower scores ($p<.001^{***}$).
For instance, the main effects of stimuli was significant on eccentricity $=60$ deg in scene \textit{bedroom} ($F(2,38)=411.93, p<.001^{***}$, \Cref{fig:lpips:bedroom}). \textbf{OURS} was significantly lower than \textbf{F-GT} ($t(19)=-3.78, p<.001^{***}$), and higher than \textbf{NeRF} ($t(19)=22.75, p<.001^{***}$) both with a ``large'' effect size (Cohen's d $>0.8$). The example observation generally applies to all 4 scenes being validated.

\paragraph{Discussion}
In the foveal and near-periphery, the observation revealed our method's significantly higher visual quality than alternative neural representation solutions. This is evidenced by the significantly stronger perceptual similarity to \textbf{F-GT} by comparison between \textbf{OURS} and \textbf{NeRF}. The subjective study (\Cref{sec:study:user}) evidenced the marginally lower similarity than \textbf{F-GT} as statistically unnoticeable among users.

In the far-periphery, \textbf{OURS} showed increased LIPIPS than \textbf{F-GT} beyond $40$ deg.
The latter has been shown to display identical perceptual similarity than full resolution rendering in VR \cite{Patney:2016:TFR}.
Thus, the findings show that \textbf{OURS} doesn't compromise the peripheral vision's quality with its significantly enhanced synthesis acuity in the fovea.

Note that the discoveries also agree with the observations from \Cref{sec:study:user}. That is, in addition to the significantly faster performance  ($99.8\%$ time reduction per frame as in \Cref{sec:study:intra}\nothing{\qisun{Check if I noted it right.}}), our method showed superior perceptual quality than \textbf{NeRF} under first-person, high resolution, and immersive viewing of 3D scenes. Meanwhile, our synthesized views are perceptually identical to the traditionally rendered images with a complete 3D mesh. It yet achieves significant data storage-saving with only about $0.5$MB neural model (up to $99.5\%$ saving from 3D mesh in our experiment).

\subsection{Performance}
\label{sec:study:intra}
% apply objective metric to |cref{fig:optimization}, aka different sample numbers
Virtual and augmented reality demands high frame rates along with high quality to ensure an immersive and comfortable experience. Our neural synthesis method introduces both spatial and angular (stereo) visual perception for real-time performance (\Cref{sec:method:representation}). 
Further, our precision-latency joint optimization for the representation and network also balances quality and performance.
Here, we evaluate the performance of each component in our system and compare with existing neural synthesis solutions.

For high resolution ($1440\times1600$), high field-of-view ($110$ deg) stereo images required by VR display, our system completes all computation (including gaze-contingent neural-inference and elemental images composition) in $31.8$ms per frame without stereo foveation, as shown in \Cref{tbl:ablation} . Compared with {\bf NeRF} (the bottom row of \Cref{tbl:ablation}), our spatial foveation significantly improves the system performance from offline computation (in seconds) to interactive speed (about 30FPS).
Our stereo foveation further reduces the computation time and achieves a high performance of more than 50 frames per second, contributing to a temporally continuous viewing experience without loss of perceived resolution and quality.

%Specifically, the significant performance gain benefits from the spatial and stereo foveation that do not compromise the visual quality.
\begin{table}[thb]
    \centering
    \begin{tabular}{ l|r } 
        \toprule
         foveal infer (per eye)  & 3.8 \\
         \midrule
         periphery infer (per eye) & 11.9 \\
          \midrule
         blending \& contrast enhancement & 0.2 \\
          \midrule\midrule
         {\bf OUR} method without stereo foveation & 31.8\\
         {\bf OUR} full method & {\bf 19.7}\\
         \midrule
         {\bf NeRF} \cite{mildenhall2020nerf} & $9\times10^3$ \\
         \bottomrule
    \end{tabular}
    \Caption{Time consumption breakdown and comparison.}
    {%
    The numbers show the average time consumption of each component and the overall system per frame. All units are in millisecond.
    \nothing{\qisun{(Jan 25) I have update the time consumption of {\bf NeRF}. See if that is correct.}}
    }
    \label{tbl:ablation}
\end{table}

### 2021-03-09 TOG/6-conclusion.tex
submission

\section{Conclusion}
\label{sec:conclusion}
\zh{consistently using high eccentricity rather far periphery?}
In this paper, we present a gaze-aware neural scene representation and view synthesis method tailored for future portable virtual reality viewing experiences. Especially, we overcome the limitations of existing neural rendering approaches for high performance, resolution, and fidelity. Our network individually synthesizes foveal, mid- and far-periphery retinal images which are then blended to form a wide field of view image matching retinal acuity. This is achieved via adapting the neural network model to the foveated visual acuity and stereopsis. 
Comparing with traditional rendering, our method requires less than $1\%$ data storage yet delivering identically high perceptual quality.
Comparing with alternative neural view synthesis approaches, our method creates significantly faster and higher fidelity viewing for high resolution, high FoV, and stereo VR head-mounted-displays.
%The experiments with user studies, objective measurements, and intra-system analysis show our significant effectiveness of delivering perceptually identical visual contents in real-time.

\paragraph{Limitations and future work}
While our method achieves unprecedented performance compared to existing approaches, it suffers from a few limitations. Our multi-spherical representation introduces higher quality when the VR camera is within the first sphere. As shown in \Cref{fig:large_translation}, when the virtual camera translates largely within a highly occluded scene, our method still synthesizes the accurate views but with a decline in foveal image quality. This is due to the accumulated error of the concentric spherical representation (\Cref{eq:imageError}).
Multiscale coordinates and networks that consider various level-of-details of the 3D space have shown their effectiveness of interpolating geometries \cite{winkler2010multi}. Developing a multiscale network that synthesizes the image from global to local level-of-details would be an interesting direction for future work, to extend the applicability of our method to large scale scenes with strong occlusions.
\begin{figure}[!ht]
    \centering
    \subfloat[small translation]{\includegraphics[width=0.47\linewidth]{TOG/figs/large_tran/2_gas_0.3.pdf}\label{fig:large_translation:small}}\hspace{1em} % or 2_gas_0.3.png
    \subfloat[large translation]{\includegraphics[width=0.47\linewidth]{TOG/figs/large_tran/2_gas_0.9.pdf}\label{fig:large_translation:large}} % or 2_gas_0.9.png
    \Caption{Comparing foveal quality with regards to translation range. }
    {%
    Using our method, \subref{fig:large_translation:small} shows a fovea image \dnc{from training dataset} with a small-scale translation box (0.3 meters); \subref{fig:large_translation:large} shows a fovea image with a large-scale translation box (0.9 meters). In the 2D illustrative inset: the blue circles indicates the distribution of the spheres; the orange arrow shows the eye's translation range.
    }
    \label{fig:large_translation}
\end{figure}

Currently, we sample the scene fully based on eccentricity, considering acuity and stereopsis. However, we envision that fine grained visual sensitivity analysis, such as luminance \cite{Tursun:2019:LCA} or depth \cite{Sun:20:OE}, would provide more insights on achieving further higher quality and/or faster performance. Also, our multi-spherical representation is specialized to the scene for optimal quality. Thus, our network is trained per-scene. Exploring potential means of generalized representation may further reduce storage.

For simplicity, the spatial-temporal joint optimization connects the output precision and latency to the number of the spheres ($\sphereNum$) but not their radii $\mathbf{\sphereRadius}$. This is due to the significant amount of training with parameter sampling. Incorporating the parameters into single training process may significantly reduce the time consumption for the optimization. With the adaptive training process, a content-aware distribution (i.e., $\mathbf{\sphereRadius}$) of the spheres would further improve the synthesis quality and performance.
%\qisun{note to myself: Right now we use did not optimize for radii but only number of spheres}

>>>>

\section{Conclusion}
\label{sec:conclusion}
In this paper, we present a gaze-aware neural scene representation and view synthesis method tailored for future portable virtual reality viewing experiences. Specifically, we overcome the limitations of existing neural rendering approaches for high performance, resolution, and fidelity. Our network individually synthesizes foveal, mid-, and far-periphery retinal images, which are then blended to form a wide field-of-view image matching retinal acuity. This is achieved via adapting the neural network model to the foveated visual acuity and stereopsis. 
Comparing with traditional rendering, our method requires less than $1\%$ data storage yet delivering identically high perceptual quality.
Comparing with alternative neural view synthesis approaches, our method creates significantly faster and higher fidelity viewing for high resolution, high FoV, and stereo VR head-mounted-displays.
%The experiments with user studies, objective measurements, and intra-system analysis show our significant effectiveness of delivering perceptually identical visual contents in real-time.

\paragraph{Limitations and future work}
While our method achieves unprecedented performance compared to existing approaches, it suffers from a few limitations. Our multi-spherical representation introduces higher quality when the VR camera is within the first sphere. As shown in \Cref{fig:large_translation}, when the virtual camera translates largely within a highly occluded scene, our method still synthesizes the accurate views but with declined foveal image quality. This is due to the accumulated error of the concentric spherical representation (\Cref{eq:imageError}).
Multiscale coordinates and networks that consider various level-of-details of the 3D space have shown their effectiveness of interpolating geometries \cite{winkler2010multi}. Developing a multiscale network that synthesizes the image from global to local level-of-details would be an interesting direction for future work, extending the applicability of our method to large-scale scenes with strong occlusions.
\begin{figure}[!ht]
    \centering
    % \subfloat[small translation]{\includegraphics[width=0.47\linewidth]{TOG/figs/large_tran/2_gas_0.3.pdf}\label{fig:large_translation:small}}\hspace{1em} % or 2_gas_0.3.png
    \subfloat[small translation]{\includegraphics[width=0.47\linewidth]{TOG/figs/large_tran/2_gas_0.3.png}\label{fig:large_translation:small}}\hspace{1em}
    % \subfloat[large translation]{\includegraphics[width=0.47\linewidth]{TOG/figs/large_tran/2_gas_0.9.pdf}\label{fig:large_translation:large}} % or 2_gas_0.9.png
    \subfloat[large translation]{\includegraphics[width=0.47\linewidth]{TOG/figs/large_tran/2_gas_0.9.png}\label{fig:large_translation:large}}
    \Caption{Comparing foveal quality with regards to translation range. }
    {%
    Using our method, \subref{fig:large_translation:small} shows a fovea image \dnc{from training dataset} with a small-scale translation box (0.3m); \subref{fig:large_translation:large} shows a fovea image with a large-scale translation box (0.9m).
    }
    \label{fig:large_translation}
\end{figure}

Currently, we sample the scene fully based on eccentricity, considering acuity and stereopsis. However, we envision that fine-grained visual sensitivity analysis, such as luminance \cite{Tursun:2019:LCA} or depth \cite{Sun:20:OE}, would provide more insights on achieving even higher quality and/or faster performance. Also, our multi-spherical representation is specialized to the scene for optimal quality. Thus, our network is trained per-scene. Exploring potential means of generalized representation may further reduce storage.

For simplicity, the spatial-temporal joint optimization in \Cref{sec:method:optimization} connects the output precision and latency to the number of the spheres ($\sphereNum$) but not their radii $\mathbf{\sphereRadius}$. This is due to the significant amount of training with parameter sampling. Incorporating the parameters into a single training process may significantly reduce the time consumption for the optimization. With the adaptive training process, a content-aware distribution (i.e., $\mathbf{\sphereRadius}$) of the spheres would further improve the synthesis quality and performance.
%\qisun{note to myself: Right now we use did not optimize for radii but only number of spheres}

\zh{shall we end with non-limitation sentences?}\qisun{yep, future work is positive than limitations.;)}
